---
title: "Task 2: Survey responses - political and environmentalist self-identification"
author: "Andrea Gracia"
date: "12/1/2020"
output: html_document
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 
In this "mini-report we explore the association between respondents’ political affiliation (“conservative” or “liberal” by self-identification) and if they consider themselves an environmentalist.The survey data explores "Influences on charitable giving for conservation.",and it was conducted in August 2017 with 1,331 yielded individual respondents.

```{r, include=FALSE, message=FALSE}
#load packages

library(tidyverse)
library(janitor)
library(here)
library(broom)
library(kableExtra)

#read in and clean data

conservation_survey <- read.csv(here("data", "conservation_survey.csv")) %>% 
  clean_names()

#subset of data that only includes relevant variables such as 'env_id' and 'pol_or'

#lines 34-42 create new columns that include political party into a binary response of either “conservative” (if pol_or is 1 - 3) or “liberal” (if pol_or 5 - 7). All other political affiliations (4, 8, 9) will be excluded.

#lines 43-47 renames outcomes as 1 = Yes, 2 = No, 3 = Unsure to the survey question "I consider myself an environmentalist"

survey_new <- conservation_survey %>% 
  select(env_id, pol_or) %>% 
  filter(pol_or %in% c(1,2,3,5,6,7)) %>% 
  mutate(pol_party = case_when(
    pol_or %in% c("1","2","3") ~ "Conservative",
    pol_or %in% c("5","6","7") ~ "Liberal")) %>% 
  mutate(env_id = case_when(
    env_id == "1" ~ "Yes",
    env_id == "2" ~ "No",
    env_id == "3" ~ "Unsure"))

#Find counts
counts_table <- survey_new %>% 
  tabyl(pol_party, env_id) 
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}

#Find proportions
count_proportions_table <- counts_table %>% 
  adorn_percentages() %>% 
  adorn_pct_formatting(digits=2) %>% 
  adorn_ns() 

kbl(count_proportions_table, 
    caption ="Table 1. Counts and proportions of “liberal” and “conservative” respondents who responded “Yes”, “No” and “Unsure” to the statement “I consider myself an environmentalist.", 
    col.names = c('Political Party','No','Unsure','Yes')) %>% 
  kable_styling()
```


```{r, include=FALSE, message=FALSE}
#Make contingency table and perform chi-square test

survey_ct<- counts_table %>% 
  column_to_rownames(var="pol_party") 

survey_x2 <- chisq.test(survey_ct)
survey_x2

survey_tidy <- tidy(survey_x2)
```

## Summary
Our findings show that there is a significant association between political association and 'environmentalist' self-identification ($\chi$^2^(`r survey_tidy$parameter`) = `r round(survey_tidy$statistic,2)`, *p* = `r round(survey_tidy$p.value, 4)`). Half of those who identified themselves as conservatives, also did not identify themselves as an environmentalist. The opposite trend is seen in the Liberal category in which 54% considered themselves an environmentalist. Overall, self-identified conservatives seem to be more certain of whether they are a conservative or not (16%) compared to liberals (21%) because of their lower percentage in the unsure category. 

## Data
Citation: Andrews Forest LTER Site and C. Batavia. 2019. Influences on charitable giving for conservation: Online survey data of 1,331 respondents across the US, August 2017 ver 3. Environmental Data Initiative.https://doi.org/10.6073/pasta/926e6270e324a1322a900da14d38b96c



