---
title: "Task 2: Comparing female lizard lenghts"
author: "Andrea Gracia"
date: "12/2/2020"
output: html_document
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 
In this mini-report we'll compare total lengths (total_length, in millimeters) for female lizards (sex == “F”) of two species (spp): CNUN (Cnemidophorus uniparens, the Desert Grassland lizard) and CNTE (Cnemidophorus tessalatus, the Colorado Checkered lizard) using a two-sample t-test and Cohen's d parameter.

```{r, include=FALSE, warning=FALSE}

#Attach packages 
library(tidyverse)
library(janitor)
library(GGally)
library(stargazer)
library(ggbeeswarm)
library(here)
library(kableExtra)
library(lubridate)
library(dplyr)
library(ggplot2)
library(effsize)
library(broom)

#Load and clean data
lizards <- read.csv(here("data", "lizards.csv")) %>% 
  clean_names()
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
### Data Visualization

#Create simplified data frame to only include CNUN and CNTE species, female and total length

#Then create extra columns that calculated mean,sd,sample size, standard error and variance of weight per sex.

lizard_length <- lizards %>% 
  select(spp,sex,total_length) %>% 
  filter(sex=="F") %>% 
  filter(spp %in% c("CNUN", "CNTE"))%>% 
  group_by(spp) %>% 
  mutate(mean_length=mean(total_length, na.rm =TRUE),
         sd_length=sd(total_length, na.rm =TRUE),
         sample_size=n(),
         se_length=sd(total_length, na.rm =TRUE)/sqrt(n()),
         var_length=var(total_length, na.rm =TRUE)) %>% 
   mutate(spp=case_when(
    spp == "CNUN" ~"Desert Grassland",
    spp == "CNTE" ~"Colorado Checkered"))
  
#This beeswarm plot compares total lengths of female lizards for the two species, while plotting mean and standard error of weight per species

ggplot()+
 geom_beeswarm(data=lizard_length,
                aes(x=spp, y=total_length, color=spp),
                size=3,
                alpha=0.5) +
   geom_point(data=lizard_length, 
             aes(x=spp, y=mean_length))+
   geom_errorbar(data=lizard_length,
                aes(x=spp,
                    ymax=mean_length + sd_length,
                    ymin=mean_length - sd_length),
                width=0.1)+
  theme_minimal()+
  labs(y="Total Length (mm)", x="Species", title="Female Colorado Checkered and Desert Grassland Lizard lengths ", color="species") 
 
```


**Figure 1** compares the means and standard errors of the two female lizard populations. The mean and standard deviation for each population is plotted, which makes it easier to visually discern how data spread compares to the average.

```{r, echo=FALSE, warning=FALSE, message=FALSE}
#Descriptive stats

#Created a summary table that contains mean and standard deviation for total length and sample size for each species

lizard_length_table <- lizards %>% 
  select(spp,sex,total_length) %>% 
  filter(sex=="F") %>% 
  filter(spp %in% c("CNUN", "CNTE"))%>% 
  group_by(spp) %>% 
  summarize(mean_length=mean(total_length, na.rm =TRUE),
         sd_length=sd(total_length, na.rm =TRUE),
         sample_size=n()) %>% 
   mutate(spp=case_when(
    spp == "CNUN" ~"Desert Grassland",
    spp == "CNTE" ~"Colorado Checkered"))
 
lizard_length_table %>% 
  kable(col.names = c("Species","Mean Length (mm)", "Standard Deviation", "Sample Size")) %>% 
  kable_styling(full_width = FALSE)
```

**Table 1** Contains the mean length (millimeters), standard deviation and sample size for female lizard populations of the Colorado Checkered and Desert Grassland populations. The table confirms findings of Figure 1, which show that Colorado Checkered lizards are longer in length on average. 

## Analysis and findings
Is there a significant difference in total length difference in mean total length between female lizards for the two species? 

```{r,include=FALSE,warning=FALSE, message=FALSE}
#2-sample t-test 

  #Exploratory data visualization
   
    #Histogram
ggplot(data=lizard_length, aes(x=total_length))+
  geom_histogram(bins=20)+
  facet_wrap(~spp)

   #Quantile-Quantile plot
ggplot(data = lizard_length, aes(sample=total_length)) +
  geom_qq()+
  facet_wrap(~spp)

# Overall: assumptions of normality hold, & sample size additionally >> Central Limit Theorem (means will be normally distributed). Two-sample t-test OK for means comparison!

```

```{r,warning=FALSE,include=FALSE, message=FALSE}

#2-sample t-test 

clean_lizard_length <- lizard_length %>% 
  select(spp,sex,total_length) %>% 
  drop_na(total_length) 

CNUN_sample <- clean_lizard_length %>% 
  filter(spp == "Desert Grassland") %>% 
  pull(total_length)

CNTE_sample <- clean_lizard_length %>% 
  filter(spp == "Colorado Checkered") %>% 
  pull(total_length)

lizard_length_ttest <-t.test(x=CNUN_sample, CNTE_sample)
lizard_length_ttest

#Cohen's d
lizard_length_effsize <- cohen.d(CNUN_sample, CNTE_sample)
lizard_length_effsize

# Get the tidy model results to call outputs in-line: 
lizard_length_ttest_tidy <- tidy(lizard_length_ttest)

# Get the means & SDs for each: 
CNUN_mean <- mean(CNUN_sample)
CNTE_mean <- mean(CNTE_sample)
CNUN_sd <- sd(CNUN_sample)
CNTE_sd <- sd(CNTE_sample)


```

On average, Colorado Checkered lizard have greater height than Desert Grassland lizards (`r round(CNUN_mean)` $\pm$ `r round(CNUN_sd)` and `r round(CNTE_mean)` $\pm$ `r round(CNTE_sd)` mm, respectively; mean $\pm$ 1 standard deviation). While the absolute difference in means is `r round(CNUN_mean) - round(CNTE_mean)` mm (a `r round((CNUN_mean - CNTE_mean)/((CNUN_mean + CNTE_mean)/2)*100)`% difference), the difference in means is significant (Welch's two-sample t-test: t(`r round(lizard_length_ttest_tidy$parameter)`) = `r round(lizard_length_ttest_tidy$statistic)`, p < 0.001), and the effect size is small (Cohen's *d* = `r round(lizard_length_effsize$estimate)`).



## Data
Lightfoot, D. 2017. Lizard pitfall trap data (LTER-II, LTER-III) ver 36. Environmental Data Initiative. https://doi.org/10.6073/pasta/ff37b4cace16a9943575f3fd7067064e 


